Predicted cardiovascular risk and blood pressure for Americans with diabetes, chronic kidney disease, and \u226565 years of age

### 2020-09-11 doc/ACCAHA_BP_groups.Rmd
cutpoints added as targets

loadd(exams)
loadd(current_analysis)
loadd(inline)
loadd(report)

@@

Data were collected during an in-home interview and a study visit completed at a mobile examination center. Standardized questionnaires were used to assess survey participants’ age, sex, race/ethnicity, smoking habits, medical history and use of antihypertensive medication, oral glucose lowering medication and insulin. Medical history included questions about whether the participant had been told by a doctor or other health professional that they had a heart attack, coronary heart disease stroke, or heart failure. Blood and urine samples were collected during the medical examination. Of relevance to the current analysis, serum creatinine, serum glucose and glycated hemoglobin (HbA1c) were measured. Diabetes was defined by fasting serum glucose ≥126 mg/dL, non-fasting glucose ≥200 mg/dL, HbA1c ≥6.5%, or self-reported use of insulin or oral glucose lowering medication. Estimated glomerular filtration rate was calculated using the Chronic Kidney Disease Epidemiology Collaboration equation.[@levey2009new] Urinary albumin and creatinine levels were measured and used to calculate the albumin-to-creatinine ratio (ACR). CKD was defined by an estimated glomerular filtration rate <60 ml/min/1.73m^2^ or an ACR ≥30 mg/dL. Ten-year predicted risk for ASCVD was calculated using the pooled cohort risk equations for participants without a history of CVD.[@goff20142013] Participants with a history of CVD were presumed to have 10-year risk for ASCVD ≥10%.

@@

@@

In the current study, the distribution of 10-year predicted risk for ASCVD was investigated among US adults aged 40 to 79 years from `r min(exams)` to `r max(exams)+1` The 2017 ACC/AHA BP guideline states that the vast majority of adults with diabetes, CKD, or ≥65 years of age have a 10-year predicted risk for ASCVD ≥10% and are recommended the initiation of antihypertensive drug therapy for stage 1 hypertension. However, the current analysis estimates that while `r inline$prevHighRiskDiabetes` of US adults with diabetes have 10-year predicted risk for ASCVD ≥10% or prevalent CVD, only `r inline$prevHighRiskS1hDiabetes` of adults with diabetes and stage 1 hypertension do. Similarly, `r inline$prevHighRiskCkd` of US adults with CKD but only `r inline$prevHighRiskS1hCkd` of US adults with CKD and stage 1 hypertension have 10-year predicted risk for ASCVD ≥10% or prevalent CVD. In contrast, among US adults with ≥65 years of age, `r inline$prevHighRiskAge65` have 10-year predicted risk for ASCVD ≥10% or prevalent CVD and `r inline$prevHighRiskS1hAge65` of US adults with stage 1 hypertension and ≥65 years of age do.

@@

The 2017 ACC/AHA BP guideline recommends initiation of antihypertensive medication based on BP level, 10-year predicted risk for ASCVD, and other conditions including diabetes, CKD, and age ≥65 years. Previous analysis of pooled data from the REasons for Geographic and Racial Differences in Stroke study and the Jackson Heart Study showed that adults with stage 1 hypertension recommended versus not recommended to initiate antihypertensive medication by the 2017 ACC/AHA BP guideline were roughly 5 times more likely to experience stroke or coronary heart disease over 12.2 years of follow-up. Moreover, 87.7% of adults with stage 1 hypertension recommended to initiate treatment by the 2017 ACC/AHA BP guideline had a 10-year predicted risk for ASCVD ≥10%. However, the mean age of adults in this group was 70.2 years, with a standard deviation of 6.9, whereas the current study found the estimated mean (95% CI) age of US adults aged 40 to 79 from 2013 to 2018 was `r inline$meanAgeOverall`. Given that age is one of the most influential predictors in the pooled cohort risk equation, the younger age of adults in the US with stage 1 hypertension recommended to initiate antihypertensive medication by the ACC/AHA BP guidelines likely explains the difference in the distribution of 10-year predicted risk for ASCVD in the current study.

@@

A prior analysis of NHANES data found that the proportion of adults in the US with SBP >130 mm Hg or DBP >80 mm Hg who are aware of their BP level declined from 2013 to 2018. Coupled with data showing that these adults may be at high risk for incident ASCVD events, it is clear that the pooled cohort risk equations or other CVD risk prediction tools should be used in clinical practice to advance critical healthcare goals such as increasing awareness of hypertension and risk for incident ASCVD. Computing predicted risk for ASCVD in the clinical setting may require additional time if a patient's medical record does not contain recent cholesterol measurements, but the current study suggests it is probably safe to assume US adults with stage 1 hypertension and ≥65 years of age have 10-year predicted risk for ASCVD ≥10%. This assumption is not supported for patients with stage 1 hypertension and diabetes or stage 1 hypertension and CKD, particularly for those aged <55 years. However, for younger adults with diabetes or CKD whose predicted 10-year risk for ASCVD is <10%, antihypertensive treatment may still provide substantial reduction in lifetime risk for ASCVD and prevention of complications associated with diabetes or CKD.

@@

This study has several limitations. First, NHANES participants completed only one visit and guidelines recommend using the mean BP aggregated over 2 or more visits. Second, the response rate for NHANES has declined form 1999-2000 to 2017-2018. The effect of this decline on the current study is unclear. Third, the use of predicted ASCVD risk rather than observed ASCVD events may have over- or under-estimated ASCVD risk in subgroups where the pooled cohort risk equations are not well-calibrated. The current study also has a number of strengths. First, NHANES data are publicly available and representative of the US population when coupled with the NHANES survey design. Second, a host of software geared to enhance the transparency and reproducibility of scientific manuscripts have been leveraged to make the current analysis entirely reproducible and open source (see first author's publicly available GitHub repository: < _Link not yet active so we won't be scooped_ >).

@@

>>>>

loadd(exams)
loadd(current_analysis)
loadd(inline)
loadd(report)
loadd(fasted_hrs_lower)
loadd(fasted_hrs_upper)
loadd(gluc_cutpoint_fasted)
loadd(gluc_cutpoint_fed)
loadd(hba1c_cutpoint)
loadd(egfr_cutpoint)
loadd(acr_cutpoint)

@@

Data were collected during an in-home interview and a study visit completed at a mobile examination center. Standardized questionnaires were used to assess survey participants’ age, sex, race/ethnicity, smoking habits, medical history and use of antihypertensive medication, oral glucose lowering medication and insulin. Medical history included questions about whether the participant had been told by a doctor or other health professional that they had a heart attack, coronary heart disease stroke, or heart failure. Blood and urine samples were collected during the medical examination. Of relevance to the current analysis, serum creatinine, serum glucose and glycated hemoglobin (HbA1c) were measured. Diabetes was defined by fasting serum glucose ≥`r gluc_cutpoint_fasted` mg/dL, non-fasting glucose ≥`r gluc_cutpoint_fed` mg/dL, HbA1c ≥`r hba1c_cutpoint`%, or self-reported use of insulin or oral glucose lowering medication. Estimated glomerular filtration rate was calculated using the Chronic Kidney Disease Epidemiology Collaboration equation.[@levey2009new] Urinary albumin and creatinine levels were measured and used to calculate the albumin-to-creatinine ratio (ACR). CKD was defined by an estimated glomerular filtration rate <`r egfr_cutpoint` ml/min/1.73m^2^ or an ACR ≥`r acr_cutpoint` mg/dL. Ten-year predicted risk for ASCVD was calculated using the pooled cohort risk equations for participants without a history of CVD.[@goff20142013] Participants with a history of CVD were presumed to have 10-year risk for ASCVD ≥10%.

@@

<!-- These are hard coded to discourage messing with guideline recs -->

@@

<!-- In the current study, the distribution of 10-year predicted risk for ASCVD was investigated among US adults aged 40 to 79 years from `r min(exams)` to `r max(exams)+1`  -->

The 2017 ACC/AHA BP guideline states that the vast majority of adults with diabetes, CKD, or ≥65 years of age have a 10-year predicted risk for ASCVD ≥10% and are recommended the initiation of antihypertensive drug therapy for stage 1 hypertension. However, the current analysis estimates that while `r inline$prevHighRiskDiabetes` of US adults with diabetes have 10-year predicted risk for ASCVD ≥10% or prevalent CVD, only `r inline$prevHighRiskS1hDiabetes` of adults with diabetes and stage 1 hypertension do. Similarly, `r inline$prevHighRiskCkd` of US adults with CKD but only `r inline$prevHighRiskS1hCkd` of US adults with CKD and stage 1 hypertension have 10-year predicted risk for ASCVD ≥10% or prevalent CVD. In contrast, among US adults with ≥65 years of age, `r inline$prevHighRiskAge65` have 10-year predicted risk for ASCVD ≥10% or prevalent CVD and `r inline$prevHighRiskS1hAge65` of US adults with stage 1 hypertension and ≥65 years of age do. Among adults whose 10-year predicted risk for ASCVD was <10%, roughly 50% of those with diabetes or CKD had 10-year predicted risk for ASCVD <5%, whereas over 90% of adults aged ≥65 years had 10-year predicted risk for ASCVD between 5% and 10%.

@@

The 2017 ACC/AHA BP guideline recommends initiation of antihypertensive medication based on BP level, 10-year predicted risk for ASCVD, and other conditions including diabetes, CKD, and age ≥65 years. Previous analysis of pooled data from the REasons for Geographic and Racial Differences in Stroke study and the Jackson Heart Study showed that adults with stage 1 hypertension recommended versus not recommended to initiate antihypertensive medication by the 2017 ACC/AHA BP guideline were roughly 5 times more likely to experience stroke or coronary heart disease over 12.2 years of follow-up. Moreover, 87.7% of adults with stage 1 hypertension recommended to initiate treatment by the 2017 ACC/AHA BP guideline had a 10-year predicted risk for ASCVD ≥10%. However, the mean age of adults in this group was 70.2 years, with a standard deviation of 6.9, whereas the estimated mean (95% CI) age of US adults with stage 1 hypertension in the current analysis was `r inline$meanAgeS1h`. Given that age is one of the most influential predictors in the pooled cohort risk equation, the younger age of adults in the US with stage 1 hypertension recommended to initiate antihypertensive medication by the ACC/AHA BP guidelines likely explains the difference in the distribution of 10-year predicted risk for ASCVD in the current study.

@@

A prior analysis of NHANES data found that the proportion of adults in the US with SBP >130 mm Hg or DBP >80 mm Hg who are aware of their BP level declined from 2013 to 2018. Coupled with data showing that these adults may be at high risk for incident ASCVD, it is clear that the pooled cohort risk equations or other CVD risk prediction tools should be used in clinical practice to advance critical healthcare goals such as increasing awareness of hypertension and controlling BP levels. Computing predicted risk for ASCVD in the clinical setting may require additional time if a patient's medical record does not contain recent cholesterol measurements, but the current study suggests it is probably safe to assume US adults with stage 1 hypertension and ≥65 years of age have 10-year predicted risk for ASCVD ≥10%. This assumption is not supported for patients with stage 1 hypertension and diabetes or stage 1 hypertension and CKD, particularly for those aged <55 years. However, for younger adults with diabetes or CKD whose predicted 10-year risk for ASCVD is <10%, antihypertensive treatment may still provide substantial reduction in lifetime risk for ASCVD and prevention of complications associated with diabetes or CKD.

@@

This study has several limitations. First, NHANES participants completed only one visit and guidelines recommend using the mean BP aggregated over 2 or more visits. Second, the response rate for NHANES has declined form 1999-2000 to 2017-2018. The effect of this decline on the current study is unclear. Third, the use of predicted ASCVD risk rather than observed ASCVD events may have over- or under-estimated ASCVD risk in subgroups where the pooled cohort risk equations are not well-calibrated. The current study also has a number of strengths. First, NHANES data are publicly available and representative of the US population when coupled with the NHANES survey design. Second, a collection of open-source software geared to enhance the transparency and reproducibility of scientific findings were leveraged to make the current analysis entirely reproducible and publicly available (see first author's GitHub repository: < _Link not yet active so we won't be scooped_ >). A point of novelty is that anyone with internet access may download this repository of code and reproduce the current manuscript as-is or with different initial parameters, e.g. using data from four NHANES cycles rather than three or modifying exclusion criteria. When different parameters are used, results in the manuscript will dynamically update. A detailed set of instructions for engaging with the project is provided in the repository.

@@

`r rp$supplement.table$caption[1]` 
`r fp_par(line_spacing = 1)`
`r rp$supplement.table$object[[1]]`
